Understanding the impacts of video-guided activities on parent-child interaction

### 2019-08-06 paper/cogsci2019/cogsci2019.Rmd
edits

---
title: "Understanding the impacts of video-guided activities on parent-child interaction"
author-information: "\\author{ \n  {\\large\\bf George~Kachergis}^1 (\\texttt{kachergis@stanford.edu})
  \\\\\n  {\\large\\bf Emily~Hembacher}^2 (\\texttt{emily.hembacher@gmail.com}) \\\\\n
  \ {\\large\\bf Veronica~Cristiano}^2 (\\texttt{CRISTIANO @gallaudet.edu}) \\\\\n
  \ {\\large\\bf Hanwen Vivian~Zhang}^1 (\\texttt{vivian3@stanford.edu}) \\\\\n  {\\large\\bf
  Michael C.~Frank}^1 (\\texttt{mcfrank@stanford.edu}) \\\\ \n^{1}Department of Psychology,
  Stanford University \\\\  Stanford, CA 94305 USA \\\\ ^{2} Nextdoor, Inc., Burlingame,
  CA \\\\ ^{3} Gallaudet University, Washington, D.C. }\n"
bibliography: library.bib
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
document-params: 10pt, letterpaper
final-submission: \cogscifinalcopy
header-includes: \usepackage{float}
keywords: "digital parenting advice; joint attention; lexical diversity; guided play;
  \n"
csl: apa6.csl
abstract: |
  Early parenting practices play an important role in shaping the future  outcomes of young children [@Hart1995; @Heckman2006].  In particular, high-quality early interactions and language input appear to  facilitate language learning and result in higher levels of school performance.  The rise of phone- and tablet-based parenting applications ("apps") holds the promise of delivering  low-cost, positive interventions on parenting style to a wide variety of populations.  Of special interest are the parents of very young children, who are often difficult to reach in other ways.  Yet little is known about the effects of communicating to parents through app-based interventions. In a study of one commercial app offering a collection of age-appropriate activity videos, we find  that the quality of parent-child interactions increases in some ways as a result of using the app.  Specifically, the lexical diversity of parents' child-directed speech increases, and measures of  joint attention show...
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, 
                      fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, 
                      message=F, sanitize = T)
```

@@

select <- dplyr::select # masked by MASS
theme_set(theme_few())
```

@@

The quantity and quality of early language input has been found to be strongly associated with later language and academic outcomes [@Hart1995; @Marchman2008; @Cartmill2013; @HirshPasek2015]. Thus, because of the potential for large downstream effects [@Heckman2006], there is tremendous interest in interventions that change children's language environment. 
And because parents define a large portion of that environment, especially before the onset of formal schooling, parent behavior is a critical locus for such interventions. 
Many effective parenting interventions require large resource investments and require many hours of in-person contact [@Jamaica2014; @PerryPreschool2004], making implementation at scale a daunting proposition. 
For this reason, many researchers targeting early language are interested in delivering parenting interventions remotely -- through texts, apps, and videos delivered on digital devices. 
But what do parents take away from these short messages about what to do or how to talk with their children?

@@

But is this theory of change correct? That is, does the provision of knowledge and activities lead to higher-quality play? 
Alternatively, this theory could be flawed in a number of ways. 
By focusing parents on a specific activity, parents may over-focus on achieving the superficial goals of the activity. 
This problem might be especially likely with video messages, which could encourage parents to try to mimic a model's specific speech and/or actions.
Attempting to reproduce such surface details of a video-guided activity could in turn result in less high-quality talk, with less responsiveness to their child's play.
Another possibility is that these messages do produce the desired effect, but only for those parents who already have a general orientation towards children's early learning.

@@

In Experiment 1, we invited parents of 6- to 24-month-old infants visiting the Children's Discovery Museum in San Jose to complete video-guided activities from Kinedu (Kinedu Inc.), a parenting app that delivers digital parenting advice in the form of short videos.
Parents were randomly assigned to the video condition or the control condition; parents in the _video_ condition watched a video from the Kinedu app (matched to their child's age), and then performed the activity with their child using the props from the video. 
Parents in the _control_ condition did not watch a Kinedu video, but were given a set of the same age-appropriate props and asked to play with their infants as they normally would at home.

@@

lmer_data <- d %>%
  #filter(!is.na(AA), !is.na(EL), !is.na(RR))%>%
  mutate(condition = factor(condition), 
         MTLD = as.numeric(MTLD),
         EL = as.numeric(langcog::scale(EL, scale=FALSE)),
         AA = as.numeric(langcog::scale(AA, scale=FALSE)),
         RR = as.numeric(langcog::scale(RR, scale=FALSE)),
         age = as.numeric(langcog::scale(age, scale=FALSE)),
         gender = as.factor(gender),
         video = as.factor(video))

@@

### Materials.
Stimuli included videos from the Kinedu commercial parenting application. 
The videos were designed to show activities to parents that they could perform with their child in order to foster cognitive and physical development, and were targeted to the child's age and level of development. 
In each video, an adult and child perform the activity (e.g., sorting toys according to size) while a narrator explains the activity and its purpose. 
We selected two videos for each of three age groups in our sample (6-11.9 months, 12-17.9 months, 18-23.94 months). 
Participants were also given a set of toys corresponding to those in the video that they watched so that they could complete the activity.[^1]

@@

Participants were randomly assigned to either the *Video* condition or the *Control* condition. 
Parents participating in the Video condition were assigned to watch one of the two activity videos available for their child's age group, while parents in the Control condition watched no activity video, and were simply asked to play with their child as they normally would.
The Control condition was yoked to the Video condition such that for every participant in the Video condition who saw a particular video and received the associated props, a participant in the Control condition received the same props but did not watch the activity video.
Parents also completed the Early Parenting Attitudes Questionnaire [EPAQ; @Hembacher2018]. 
The EPAQ measures parents' attitudes about parenting and child development along three dimensions: rules and respect, early learning, and affection and attachment.

### Procedure.
After providing informed consent, parents in the Video condition watched the assigned activity video on a laptop with headphones.
To ensure that parents could give the video their full attention, the experimenter played with the infant with a set of toys (different from the experimental props used in the study) while the video was being played. 
Immediately following the video, each parent-child dyad was provided with the props to complete the video-guided activity that the parent had viewed. 
The toys were placed on a large foam play mat, and parents were instructed to sit on the mat with their child and re-create the activity they had viewed for a period of three minutes.
In the Control condition, after informed consent parents were told to play with their child as they would at home with the provided props for a period of three minutes. 
They were not given any additional instructions about how to use the props.

In both conditions, two video cameras were used to record the play session from different angles, and parents were fitted with a wireless Shurre lavalier microphone to record their child-directed speech. 
After three minutes of play had elapsed, parents were told they could stop playing and the cameras and microphone were turned off.
Parents were then asked to complete the EPAQ before being debriefed.

### Joint Attention Coding Procedure.
The video of each session was manually coded for episodes of joint attention (JA) using the Datavyu software [@datavyu]. 
The video taken at floor level was coded by default, but the other video was referred to if the participants were not visible or if there was technical difficulty with the first camera. 
Each session's video was coded for episodes of coordinated JA, episodes of passive JA, and parental bids for JA. 
Parental bids for JA were defined as any attempt to initiate joint attention (i.e labeling, pointing, or otherwise drawing attention to an object) that did not result in passive or coordinated JA. 
If more than 3 seconds elapsed between bids, they were coded as separate attempts. 
An episode of joint attention was considered _passive_ if both participants visually focused on an object for 3 or more seconds but the child did not acknowledge the parent. 
If either participant looked away from the object for less than 3 seconds and then returned to the same object it was considered part of the same episode of joint attention. 
A joint attention episode was considered _coordinated_ if both participants visually focused on an object for 3 or more seconds and at some point in the interaction the child indicated awareness of interaction with some overt behavior toward the parent such as looking at their face, gesturing, vocalizing, or turn-taking. 
Full details of our guidelines for coding joint attention are available [here](https://docs.google.com/document/d/1cZrT_Gjt6p0om19lXO-XcFnT09pb0giyhRQ-Pm1yoBU/edit?usp=sharing).

@@

Parents' child-directed speech during the play sessions was transcribed.
The transcripts and hand-coded joint attention data were analyzed according to our preregistration[^2], with any deviations or extensions noted.
Below we first report the lexical diversity results, followed by the joint attention results.

[^2]: Preregistration: [https://osf.io/2bpdf/](https://osf.io/2bpdf/)]

@@

require("lmerTest")
# Predicting lexical diversity based on experimental condition, PAQ, demographics. 
modTTR <- lmer(TTR ~ condition + age + gender + parent_ed + (1|video), data=lmer_data)
summary(modTTR) # condition:Exp -0.113 t(52.4) = 4.22 p<.001
e1TTR = get_lmer_reporting_values(modTTR)

modMTLD <- lmer(MTLD ~ condition + age + gender + parent_ed + (1|video), data=lmer_data)
summary(modMTLD) # condition -8.73 t(55) = 2.67 p=.001 !! singular fit
e1MTLD = get_lmer_reporting_values(modMTLD)
# Bayesian regression (stan_glmer in rstanarm) shows the same:
require("rstanarm")
bmodMTLD <- stan_glmer(MTLD ~ condition + age + gender + parent_ed + (1|video), data=lmer_data)
summary(bmodMTLD) # condition mean=-8.6 sd=3.2 95% CI = [-14.9,-2.3]

@@

# predict word types
types_mod <- lmer(types ~ condition*EL + condition*AA + condition*RR  + age + gender + parent_ed + (1|video), data=lmer_data) # !! singular fit
summary(types_mod) # nothing (hint: conditionexp:AA beta=19.04 p=.11)
e1types = get_lmer_reporting_values(types_mod) 
# verify with Bayesian regression:
bmod_types <- stan_glmer(types ~ condition*EL + condition*AA + condition*RR  + age + gender + parent_ed + (1|video), data=lmer_data)
summary(bmod_types) # nothing

@@

We fit a mixed-effects linear regression predicting TTR as a function of condition, age (0-centered), gender, and parent's education level with a random intercept per video using lme4 [@lme4]. 
There was significantly lower TTR in the Video condition (mean: `r round(subset(e1_lexdiv, condition=='exp')$TTR, 2)`) than in the Control condition (mean: `r round(subset(e1_lexdiv, condition=='con')$TTR, 2)`, $\beta=$`r e1TTR["conditionexp","beta"]`, t(52.4) = `r e1TTR["conditionexp","tval"]`, *p*<.001). 
There were no significant effects of age, gender, or parent's level of education.
A similar mixed-effects linear regression instead predicting MTLD also found significantly lower lexical diversity in the Video condition (mean MTLD: `r round(subset(e1_lexdiv, condition=='exp')$MTLD, 2)`) than in the Control condition (mean: `r round(subset(e1_lexdiv, condition=='exp')$TTR, 2)`, $\beta=$`r e1MTLD["conditionexp","beta"]`, t(55) = `r e1MTLD["conditionexp","tval"]`, *p*=.001), with no other significant effects.
Figure \ref{fig:e1lex_div} shows the mean of each lexical diversity measure (TTR and MTLD) by condition.

```{r e1tokens, echo=F, include=F}
d$Condition = factor(d$condition, levels = c("con", "exp"), labels = c("Control", "Video"))

@@

We also conducted similar regressions predicting the number of word tokens and types, finding only a significant effect of condition on the number of word tokens ($\beta= `r e1tokens["conditionexp","beta"]`$, t(40.4)=`r e1tokens["conditionexp","tval"]`, *p*<.05), with parents using more words in the Video condition (mean: `r round(subset(e1_lexdiv, condition=='exp')$tokens, 0)`, bootstrapped 95% confidence intervals (CI): [`r round(subset(ms_tok, Condition=="Video")$ci_lower, 0)`,`r round(subset(ms_tok, Condition=="Video")$ci_upper, 0)`]) than in the Control condition (mean: `r round(subset(e1_lexdiv, condition=='con')$tokens, 0)`, bootstrapped 95% CI: [`r round(subset(ms_tok, Condition=="Control")$ci_lower, 0)`,`r round(subset(ms_tok, Condition=="Control")$ci_upper, 0)`]).
The means of the lexical diversity measures are reported in Table \ref{e1tab}.

@@

```{r fig:e1lex_div, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3.5, fig.height=3.0, set.cap.width=T, num.cols.cap=1, fig.cap = "\\label{fig:e1lex_div} Mean lexical diversity scores by condition (left: Type/Token ratio, right: MTLD) in Experiment 1. Both measures of lexical diversity were significantly lower after watching an activity video. Error bars show bootstrapped 95\\% confidence intervals (CIs), and gray dots indicate values for each participant."}

@@

multiplot(e1ttr, e1mtld, cols=2)
```

Word tokens and word types

```{r e1token_type, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3.5, fig.height=3.0, set.cap.width=T, num.cols.cap=1, fig.cap = "Mean number of word tokens (left) and word types (right) by condition in Experiment 1. Parents used significantly more word tokens after watching an activity video. Error bars show bootstrapped 95\\% CIs."}

@@

e1types <- ggplot(ms_type, aes(x = Condition, y = mean, fill = Condition)) + geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width = .9)) +
  geom_point(data=d, aes(x=Condition, y=types), alpha=.2, position = position_jitterdodge()) +
  xlab("Condition") + ylab("Total Number of Word Types") +
  langcog::scale_fill_solarized() + ggthemes::theme_few() +
  theme(legend.position="none")

multiplot(e1tokens, e1types, cols=2)
```

@@

```{r, e1ja-regressions, echo=F, include=F}
load(paste(e1path,"joint_attention/Exp1_joint_attention_data.RData",sep=''))
d$Condition <- factor(d$Condition, levels = c("con","exp"), labels = c("Control","Video"))

@@

# Total duration of coordinated joint attention.
cja_time_mod <- lmer(paste("cja_length", model_string), data = lmer_data)
summary(cja_time_mod) # marginal parent education p=.09 - SINGULAR (do stan_glmer)
```

We fit a mixed-effects linear regression predicting the number of bids for joint attention (JA) as a function of fixed effects of condition, age (scaled and 0-centered), gender, parent's education level, and the subscales of the EPAQ: Early Learning (EL), Affection and Attachment (AA), and Rules and Respect (RR), along with interactions of condition and EL, AA, and RR. 
This lme4 model included random intercepts per video. 
There were significantly more bids for JA in the Video condition (mean: `r round(subset(e1bids_M, Condition=="Video")$mean,2)`, sd: `r round(subset(e1bids_M, Condition=="Video")$sd,2)`) than in the Control condition (mean: `r round(subset(e1bids_M, Condition=="Control")$mean,2)`, sd: `r round(subset(e1bids_M, Condition=="Control")$sd,2)`, $\beta=`r e1bids["ConditionVideo","beta"]`$, t(`r e1bids["ConditionVideo","df"]`) = `r e1bids["ConditionVideo","tval"]`, *p*<.01). 
There were no other significant effects.

Mixed-effects regressions with the same structure were performed predicting the number of episodes of coordinated and passive JA, and the total duration of time spent in coordinated and passive JA.
There were no significant effects on the number or total duration of coordinated JA episodes, nor on the total duration of passive JA episodes. 
For the regression predicting the number of passive JA episodes, the only significant effect was an interaction of condition and RR ($\beta=`r e1pja["ConditionVideo:RR","beta"]`$, t(`r e1pja["ConditionVideo:RR","df"]`) = `r e1pja["ConditionVideo:RR","tval"]`, *p*<.05), showing that for parents in the Video condition, those with higher Rules and Respect subscores engaged in more passive JA episodes.
Figure \ref{fig:e1ja} shows the mean number bids for JA and episodes of JA by condition in Experiment 1.

```{r fig:e1ja, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3.5, fig.height=3.0, set.cap.width=T, num.cols.cap=1, fig.cap = "\\label{fig:e1ja} Mean number of bids (left) and episodes (right) of joint attention (JA) by condition in Experiment 1. Parents made significantly more bids for JA after watching an activity video, but this did not result in a greater number of episodes of JA."}

ms_bids <- d %>% group_by(Condition) %>%
  tidyboot_mean(bids_tot)

e1bids <- ggplot(ms_bids, aes(x = Condition, y = mean, fill = Condition)) + geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width = .9)) +
  geom_point(data=d, aes(x=Condition, y=bids_tot), alpha=.2, position = position_jitterdodge()) +
  xlab("Condition") + ylab("Mean Bids for Joint Attention") +
  langcog::scale_fill_solarized() + ggthemes::theme_few() + 
  theme(legend.position="none")

# total joint attention episodes
ms_tja <- d %>% group_by(Condition) %>%
  tidyboot_mean(total_ja)

e1tja <- ggplot(ms_tja, aes(x = Condition, y = mean, fill = Condition)) + geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width = .9)) +
  geom_point(data=d, aes(x=Condition, y=total_ja), alpha=.2, position = position_jitterdodge()) +
  xlab("Condition") + ylab("Mean Episodes of Joint Attention") +
  langcog::scale_fill_solarized() + ggthemes::theme_few() + 
  theme(legend.position="none")

@@

```{r fig:e1ja-pass-coord, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3.5, fig.height=3.5, set.cap.width=T, num.cols.cap=1, fig.cap = "\\label{fig:e1ja-pass-coord} The mean number of passive (left) and coordinated episodes (right) of JA  in Experiment 1 were similar in both conditions."}
ms_pja <- d %>% group_by(Condition) %>%
  tidyboot_mean(pja)

e1pja <- ggplot(ms_pja, aes(x = Condition, y = mean, fill = Condition)) + geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width = .9)) +
  geom_point(data=d, aes(x=Condition, y=pja), alpha=.2, position = position_jitterdodge()) +
  xlab("Condition") + ylab("Episodes of Passive Joint Attention") +
  langcog::scale_fill_solarized() + ggthemes::theme_few() + theme(legend.position="none")

>>>>

---
title: "Understanding the impacts of video-guided activities on parent-child interaction"
author-information: "\\author{ \n  {\\large\\bf George~Kachergis}^1 (\\texttt{kachergis@stanford.edu})
  \\\\\n  {\\large\\bf Emily~Hembacher}^2 (\\texttt{emily.hembacher@gmail.com}) \\\\\n
  \ {\\large\\bf Veronica~Cristiano}^2 (\\texttt{CRISTIANO @gallaudet.edu}) \\\\\n
  \ {\\large\\bf Hanwen Vivian~Zhang}^1 (\\texttt{vivian3@stanford.edu}) \\\\\n  {\\large\\bf
  Michael C.~Frank}^1 (\\texttt{mcfrank@stanford.edu}) \\\\ \n^{1}Department of Psychology,
  Stanford University \\\\  Stanford, CA 94305 USA \\\\ ^{2} Nextdoor, Inc., Burlingame,
  CA \\\\ ^{3} Gallaudet University, Washington, D.C. }\n"
bibliography: library.bib
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
document-params: 10pt, letterpaper
final-submission: \cogscifinalcopy
header-includes: \usepackage{float}
keywords: "digital parenting advice; joint attention; lexical diversity; guided play;
  \n"
csl: apa6.csl
abstract: |
  Early parenting practices play an important role in shaping the future  outcomes of young children [@Hart1995; @Heckman2006].  In particular, high-quality early interactions and language input appear to  facilitate language learning and result in higher levels of school performance. The rise of phone- and tablet-based parenting applications ("apps") holds the promise of delivering  low-cost, positive interventions on parenting style to a wide variety of populations. Of special interest are the parents of very young children, who are often difficult to reach in other ways. Yet little is known about the effects of communicating to parents through app-based interventions. We showed parents a short video depicting an age-appropriate parent-child activity from a commercial parenting app, and found that the quality of parent-child interactions increases in some ways as a result of the intervention. Specifically, after watching the activity video, parents spoke more and made more bids for joint attention with the child.
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=4, fig.height=3, fig.crop = F, 
                      fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, 
                      message=F, sanitize = T)
```

@@

select <- dplyr::select # masked by MASS
theme_set(theme_few())

# labels=addline_format(c("Activity Video", "Science Video"))
addline_format <- function(x,...){
    gsub('\\s','\n',x)
}
```

@@

The quantity and quality of early language input has been found to be strongly associated with later language and academic outcomes [@Hart1995; @Marchman2008; @Cartmill2013; @HirshPasek2015]. Thus, because of the potential for large downstream effects [@Heckman2006], there is tremendous interest in interventions that change children's language environment. 
And because parents define a large portion of that environment, especially before the onset of formal schooling, parent behavior is a critical locus for such interventions. 
Many effective parenting interventions require large resource investments and require many hours of in-person contact [@Jamaica2014; @PerryPreschool2004], making implementation at scale a daunting proposition. 
For this reason, many researchers targeting early language are interested in delivering parenting interventions remotely -- through texts, apps, and videos delivered on digital devices. 
But what do parents take away from these short messages about what to do with or how to talk with their children?

@@

But is this theory of change correct? That is, does the provision of knowledge and activities lead to higher-quality play? 
Alternatively, by focusing parents on a specific activity, this theory could be flawed, causing parents to over-focus on achieving the superficial goals of the activity. 
This problem might be especially likely with video messages, which could encourage parents to try to mimic a model's specific speech and/or actions.
Attempting to reproduce such surface details of a video-guided activity could in turn result in less high-quality talk, with less responsiveness to their child's play.
Another possibility is that these messages might produce the desired effect, but only for those parents who already have a general orientation towards children's early learning.

@@

In Experiment 1, we invited parents of 6- to 24-month-old infants visiting the Children's Discovery Museum in San Jose to complete video-guided activities from a commerical parenting app that delivers digital parenting advice in the form of short videos.
Parents were randomly assigned to the video condition or the control condition; parents in the _activity video_ condition watched a video from the app (matched to their child's age), and then performed the activity with their child using the props from the video. 
Parents in the _control_ condition did not watch an activity video, but were given a set of the same age-appropriate props and asked to play with their infants as they normally would at home.

@@

lmer_data <- d %>%
  #filter(!is.na(AA), !is.na(EL), !is.na(RR))%>%
  mutate(condition = factor(condition), 
         MTLD = as.numeric(MTLD),
         EL = as.numeric(langcog::scale(EL, scale=FALSE)),
         AA = as.numeric(langcog::scale(AA, scale=FALSE)),
         RR = as.numeric(langcog::scale(RR, scale=FALSE)),
         age = as.numeric(langcog::scale(age, scale=FALSE)),
         parent_ed = as.numeric(langcog::scale(parent_ed, scale=T)),
         gender = as.factor(gender),
         video = as.factor(video))

@@

### Materials.
Stimuli included activity videos from a commercial parenting application. 
The videos were designed to show activities to parents that they could perform with their child in order to foster cognitive and physical development, and were targeted to the child's age and level of development. 
In each video, an adult and child perform the activity (e.g., sorting toys according to size) while a narrator explains the activity and its purpose. 
We selected two videos for each of three age groups in our sample (6-11.9 months, 12-17.9 months, 18-23.94 months). 
Participants were also given a set of toys corresponding to those in the video that they watched so that they could complete the activity.[^1]

@@

Participants were randomly assigned to either the *Activity Video* condition or the *No Video* condition. 
Parents participating in the Activity Video condition were assigned to watch one of the two activity videos available for their child's age group, while parents in the No Video condition watched no video, and were simply asked to play with their child as they normally would.
The No Video condition was yoked to the Activity Video condition such that for every participant in the Video condition who saw a particular video and received the associated props, a participant in the No Video condition received the same props but did not watch the activity video.
Parents also completed the Early Parenting Attitudes Questionnaire [EPAQ; @Hembacher2018]. 
The EPAQ measures parents of young children's attitudes about parenting and child development along three dimensions: rules and respect, early learning, and affection and attachment.

### Procedure.
After providing informed consent, parents in the Activity Video condition watched the assigned activity video on a laptop with headphones.
To ensure that parents could give the video their full attention, the experimenter played with the infant with a set of toys (different from the experimental props used in the study) while the video was being played. 
Immediately following the video, each parent-child dyad was provided with the props to complete the video-guided activity that the parent had viewed. 
The toys were placed on a large foam play mat, and parents were instructed to sit on the mat with their child and re-create the activity they had viewed for a period of three minutes.[^2]
In the Control condition, after informed consent parents were told to play with their child as they would at home with the provided props for a period of three minutes. 
They were not given any additional instructions about how to use the props.

[^2]: Based on piloting, we estimated these activities would would only require three minutes to complete.

In both conditions, two video cameras were used to record the play session from different angles, and parents were fitted with a wireless Shure lavalier microphone to record their child-directed speech. 
After three minutes of play had elapsed, parents were told they could stop playing and the cameras and microphone were turned off.
Parents were then asked to complete the EPAQ before being debriefed.

### Joint Attention Coding Procedure.
The video of each session was manually coded for episodes of joint attention (JA) using the Datavyu software [@datavyu]. 
The video taken at floor level was coded by default, but the other video was referred to if the participants were occluded or if there was technical difficulty with the first camera. 
Each session's video was coded for episodes of coordinated JA, episodes of passive JA, and parental bids for JA. 
Parental bids for JA were defined as any attempt to initiate joint attention (i.e labeling, pointing, or otherwise drawing attention to an object) that did not result in passive or coordinated JA. 
If more than 3 seconds elapsed between bids, they were coded as separate attempts. 
An episode of joint attention was considered _passive_ if both participants visually focused on an object for 3 or more seconds but the child did not acknowledge the parent. 
If either participant looked away from the object for less than 3 seconds and then returned to the same object it was considered part of the same episode of joint attention. 
A joint attention episode was considered _coordinated_ if both participants visually focused on an object for 3 or more seconds and at some point in the interaction the child indicated awareness of interaction with some overt behavior toward the parent such as looking at their face, gesturing, vocalizing, or turn-taking. 
Full details of our guidelines for coding joint attention are available in [SI](https://docs.google.com/document/d/1cZrT_Gjt6p0om19lXO-XcFnT09pb0giyhRQ-Pm1yoBU/edit?usp=sharing).

@@

Parents' child-directed speech during the play sessions was transcribed.
The transcripts and hand-coded joint attention data were analyzed according to our preregistration[^3], with any deviations or extensions noted.
Below we first report the lexical diversity results, followed by the joint attention results.

[^3]: Preregistration: [https://osf.io/2bpdf/](https://osf.io/2bpdf/)]

@@

require("lmerTest")

# Predicting lexical diversity based on experimental condition, PAQ, demographics. 
modTTR <- lmer(TTR ~ condition + age + gender + parent_ed + (1|video), data=lmer_data)
summary(modTTR) # condition:Exp -0.113 t(52.4) = 4.22 p<.001
e1TTR = get_lmer_reporting_values(modTTR)

modMTLD <- lmer(MTLD ~ condition + age + gender + parent_ed + (1|video), data=lmer_data)
 # , control=lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))
summary(modMTLD) # condition -8.73 t(55) = 2.67 p=.001 !! singular fit
e1MTLD = get_lmer_reporting_values(modMTLD)

tt <- getME(modMTLD,"theta")
ll <- getME(modMTLD,"lower")
min(tt[ll==0]) # 0

# Bayesian regression (stan_glmer in rstanarm) shows the same:
require("rstanarm")
bmodMTLD <- stan_glmer(MTLD ~ condition + age + gender + parent_ed + (1|video), data=lmer_data)
summary(bmodMTLD) # condition mean=-8.6 sd=3.2 95% CI = [-14.9,-2.3]

@@

# predict word types
types_mod <- lmer(types ~ condition*EL + condition*AA + condition*RR  + age + gender + parent_ed + (1|video), data=lmer_data,
                  control=lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5))) # !! singular fit
summary(types_mod) # nothing (hint: conditionexp:AA beta=19.04 p=.11)
e1types = get_lmer_reporting_values(types_mod) 
# verify with Bayesian regression:
bmod_types <- stan_glmer(types ~ condition*EL + condition*AA + condition*RR  + age + gender + parent_ed + (1|video), data=lmer_data)
summary(bmod_types) # nothing

@@

We fit a mixed-effects linear regression predicting TTR as a function of condition, age (centered), gender, and parent's education level with a random intercept per video using lme4 [@lme4]. 
There was significantly lower TTR in the Video condition (mean: `r round(subset(e1_lexdiv, condition=='exp')$TTR, 2)`) than in the Control condition (mean: `r round(subset(e1_lexdiv, condition=='con')$TTR, 2)`, $\beta=$`r e1TTR["conditionexp","beta"]`, t(52.4) = `r e1TTR["conditionexp","tval"]`, *p*<.001). 
There were no significant effects of age, gender, or parent's level of education.
A similar mixed-effects linear regression instead predicting MTLD also found significantly lower lexical diversity in the Video condition (mean MTLD: `r round(subset(e1_lexdiv, condition=='exp')$MTLD, 2)`) than in the Control condition (mean: `r round(subset(e1_lexdiv, condition=='exp')$TTR, 2)`, $\beta=$`r e1MTLD["conditionexp","beta"]`, t(55) = `r e1MTLD["conditionexp","tval"]`, *p*=.001), with no other significant effects.
Figure \ref{fig:e1lex_div} shows the mean of each lexical diversity measure (TTR and MTLD) by condition.

```{r e1tokens, echo=F, include=F}
d$Condition = factor(d$condition, levels = c("con", "exp"), labels = c("No Video", "Activity Video"))

@@

We also conducted similar regressions predicting the number of word tokens and types, finding only a significant effect of condition on the number of word tokens ($\beta= `r e1tokens["conditionexp","beta"]`$, t(40.4)=`r e1tokens["conditionexp","tval"]`, *p*<.05), with parents using more words in the Video condition (mean: `r round(subset(e1_lexdiv, condition=='exp')$tokens, 0)`, bootstrapped 95% confidence intervals (CI): [`r round(subset(ms_tok, Condition=="Activity Video")$ci_lower, 0)`,`r round(subset(ms_tok, Condition=="Activity Video")$ci_upper, 0)`]) than in the Control condition (mean: `r round(subset(e1_lexdiv, condition=='con')$tokens, 0)`, bootstrapped 95% CI: [`r round(subset(ms_tok, Condition=="No Video")$ci_lower, 0)`,`r round(subset(ms_tok, Condition=="No Video")$ci_upper, 0)`]).
The means of the lexical diversity measures are reported in Table \ref{e1tab}.

@@

```{r fig:e1lex_div, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=7, fig.height=3.0, set.cap.width=T, num.cols.cap=1, fig.cap = "\\label{fig:e1lex_div} Mean lexical diversity scores by condition (left: Type/Token ratio, right: MTLD) in Experiment 1. Both measures of lexical diversity were significantly lower after watching an activity video. Error bars show bootstrapped 95\\% confidence intervals (CIs), and gray dots indicate values for each participant."}

@@

@@

multiplot(e1ttr, e1mtld, e1tokens, cols=3)
```

@@

```{r, e1ja-regressions, echo=F, include=F}
load(paste(e1path,"joint_attention/Exp1_joint_attention_data.RData",sep=''))
d$Condition <- factor(d$Condition, levels = c("con","exp"), labels = c("No Video","Activity Video"))

@@

# Total duration of coordinated joint attention.
cja_time_mod <- lmer(paste("cja_length", model_string), data = lmer_data,
                     control=lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))
summary(cja_time_mod) # marginal parent education p=.09 - SINGULAR (do stan_glmer)
```

We fit a mixed-effects linear regression predicting the number of bids for joint attention (JA) as a function of fixed effects of condition, age (centered), gender, parent's education level, and the subscales of the EPAQ: Early Learning (EL), Affection and Attachment (AA), and Rules and Respect (RR), along with interactions of condition and EL, AA, and RR. 
This lme4 model included random intercepts per video. 
There were significantly more bids for JA in the Video condition (mean: `r round(subset(e1bids_M, Condition=="Activity Video")$mean,2)`, sd: `r round(subset(e1bids_M, Condition=="Activity Video")$sd,2)`) than in the Control condition (mean: `r round(subset(e1bids_M, Condition=="No Video")$mean,2)`, sd: `r round(subset(e1bids_M, Condition=="No Video")$sd,2)`, $\beta=`r e1bids["ConditionActivity Video","beta"]`$, t(`r e1bids["ConditionActivity Video","df"]`) = `r e1bids["ConditionActivity Video","tval"]`, *p*<.01). 
There were no other significant effects.

Mixed-effects regressions with the same structure were performed predicting the number of episodes of coordinated and passive JA, and the total duration of time spent in coordinated and passive JA.
There were no significant effects on the number or total duration of coordinated JA episodes, nor on the total duration of passive JA episodes. 
For the regression predicting the number of passive JA episodes, the only significant effect was an interaction of condition and RR ($\beta=`r e1pja["ConditionActivity Video:RR","beta"]`$, t(`r e1pja["ConditionActivity Video:RR","df"]`) = `r e1pja["ConditionActivity Video:RR","tval"]`, *p*<.05), showing that for parents in the Video condition, those with higher Rules and Respect subscores engaged in more passive JA episodes.
Figure \ref{fig:e1ja} shows the mean number bids for JA and episodes of JA by condition in Experiment 1.

```{r fig:e1ja, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=5.5, fig.height=3.0, set.cap.width=T, num.cols.cap=1, fig.cap = "\\label{fig:e1ja} Mean number of bids (left) and episodes (right) of joint attention (JA) by condition in Experiment 1. Parents made significantly more bids for JA after watching an activity video, but this did not result in a greater number of episodes of JA."}

ms_bids1 <- d %>% group_by(Condition) %>%
  tidyboot_mean(bids_tot)

e1bids <- ggplot(ms_bids1, aes(x = Condition, y = mean, fill = Condition)) + geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width = .9)) +
  geom_point(data=d, aes(x=Condition, y=bids_tot), alpha=.2, position = position_jitterdodge()) +
  xlab("Condition") + ylab("Mean Bids for Joint Attention") +
  langcog::scale_fill_solarized() + ggthemes::theme_few() + 
  theme(legend.position="none")

# total joint attention episodes
ms_tja1 <- d %>% group_by(Condition) %>%
  tidyboot_mean(total_ja)

e1tja <- ggplot(ms_tja1, aes(x = Condition, y = mean, fill = Condition)) + geom_bar(stat="identity") + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width = .9)) +
  geom_point(data=d, aes(x=Condition, y=total_ja), alpha=.2, position = position_jitterdodge()) +
  xlab("Condition") + ylab("Mean Episodes of Joint Attention") +
  langcog::scale_fill_solarized() + ggthemes::theme_few() + 
  theme(legend.position="none")

@@
